---
title: "SVR-with-e1071"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SVR-with-e1071}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(e1071)
library(tidyquant)
```


# Importing BTC and NVDA stock data

```{r, cache = T}
tickers = c("BTC-USD", "ETH-USD")

prices <- tq_get(tickers, from = "2018-01-01")
prices_adj = prices[,c("symbol", "date", "adjusted")]
library(tidyr)
prices_adj_wide =
  prices_adj %>%
  pivot_wider(names_from = symbol, values_from = adjusted)
```


## Plotting data

```{r, fig.width= 15}
library(ggplot2)
prices %>%
  ggplot(aes(x = date, y = adjusted, color = symbol)) +
  geom_line() +
  facet_wrap(~symbol,scales = 'free_y') +
  theme_classic() +
  labs(x = 'Date',
       y = "Adjusted Price",
       title = "Price Chart") +
  scale_x_date(date_breaks = "year",
               date_labels = "%b\n%y")
```


## Fitting models

The `e1071::svm()` function implements SVR, epsilon regression, when the outcome is not a factor. The default value of epsilon is set to 0.1. In the paper they use zero.

This is using the same end of day ETH price to predict the BTC price 

```{r}
model_svm = e1071::svm(data = prices_adj_wide, `BTC-USD` ~ `ETH-USD`)
```

```{r}
predYsvm = predict(model_svm)
```

Can see from plotting data against the predictions that we nee to carry out some model tuning
```{r}
plot(y = prices_adj_wide$`BTC-USD`, x = prices_adj_wide$`ETH-USD`)
points(prices_adj_wide$`BTC-USD`, predYsvm, col = "red", pch=16)
```


### Tuning 

```{r}
tune.svm(svm, `BTC-USD` ~ `ETH-USD`, data = prices_adj_wide)
```




```{r}
library(quantmod)
library(tidyquant)

getSymbols("BTC-USD", from = '2018-08-01',
           to = "2020-12-31",warnings = FALSE,
           auto.assign = TRUE)
```


### Using `import_stonks()`

#### Trying Tuning

```{r}
data = import_stonks()
model_svm = e1071::svm(data = data, `BTC_USD` ~ .)
```

```{r}
predYsvm = predict(model_svm)
```

```{r, cache = T}
data = import_stonks(day_lag = c(1,2,3))


gamma_range=2^(-1:1)
C_range=2^(0:4)
eps_range=0.1*(0:5)

tuneSVM = tune.svm(BTC_USD~., data = data,  sampling = "cross", type = "eps-regression", kernel = "radial",
                   gamma= gamma_range, cost=C_range, epsilon= eps_range)

tuneSVM$best.parameters
tuneSVM$
```

Best parameters: gamma = 1, cost = 2, epsilon = 0.1


```{r}
plot(y = predict(tuneSVM$best.model), 1:nrow(data))
points(y = data$BTC_USD, 1:nrow(data), col = "red")
```


### Tuning with glasso selected stocks

```{r}
tickers = c("BTC-USD","ETH-USD","BKNG","AZO","CMG","MTD","NVR")
```

```{r}
data = import_stonks(stock_pred = tickers)
model_svm = e1071::svm(data = data, `BTC_USD` ~ .)
```

```{r}
predYsvm = predict(model_svm)
```

```{r, cache = T}
data = import_stonks(day_lag = c(1,2,3))


gamma_range=2^(-1:1)
C_range=2^(0:4)
eps_range=0.1*(0:5)

tuneSVM = tune.svm(BTC_USD~., data = data,  sampling = "cross", type = "eps-regression", kernel = "radial",
                   gamma= gamma_range, cost=C_range, epsilon= eps_range)

tuneSVM$best.parameters
tuneSVM$
```

Best parameters: gamma = 1, cost = 2, epsilon = 0.1


```{r}
plot(y = predict(tuneSVM$best.model), 1:nrow(data))
points(y = data$BTC_USD, 1:nrow(data), col = "red")
```
