---
title: "testing-glasso"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{testing-glasso}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
#library(SVMForecast)
library(tidyquant)
library(ggplot2)
```


## Importing SP500

We can extract the stock tickers for the S&P 500 using `tq_index()`. The stocks are then listed in descending order with respect to their weighting in the index. 
```{r}
SP500 = tq_index("SP500")
SP500_symbols = SP500$symbol
```

Can then use this to import the stock data for each of the stocks in the S&P 500. 

```{r}
SP500_prices <- tq_get(SP500_symbols, from = "2020-01-01", to = "2022-01-01")
SP500_prices <- tq_get(SP500_symbols[1:100], from = "2020-01-01", to = "2021-01-01")
SP500_prices_adj = SP500_prices[,c("symbol", "date", "adjusted")]
```

```{r}
library(tidyr)
SP500_prices_adj_wide =
  SP500_prices_adj %>%
  pivot_wider(names_from = symbol, values_from = adjusted)
```




## glasso

```{r}
library(glasso)
```

```{r}
SP500_prices_adj_wide_mat = SP500_prices_adj_wide[,-1]
SP500_prices_adj_wide_mat = as.matrix(SP500_prices_adj_wide_mat)
cov_mat = cov(SP500_prices_adj_wide_mat)
```


Running `glasso`

```{r}
glasso = glasso(cov_mat, rho = 3000)
glasso_w = glasso$w
```

Now want to plot the network, using the precision matrix as the adjacency matrix.

```{r}
library(igraph)
```

```{r}
diag(glasso_w) = 0
```

```{r}
glasso_graph = graph_from_adjacency_matrix(glasso_w, mode = "undirected")
isolated <-  which(degree(glasso_graph) == 0)
glasso_graph <- delete.vertices(glasso_graph, isolated)
```

```{r}
plot(glasso_graph)
```


# Using glasso to choose set of predictors

```{r}
tickers = unique(c("BTC-USD", "ETH-USD", "BNB-USD", "USDT-USD", "ADA-USD", "DOGE-USD", "XRP-USD", "LTC-USD", "TSLA", "NVDA", "AMD", "PYPL", SP500_symbols))

#SOL-USD
# UNI-USD
# DOT-USD

data = tidyquant::tq_get(tickers, from = "2019-01-01")
data = data[,c("symbol", "date", "adjusted")]
```

```{r}
library(tidyr)
data_wide =
  data %>%
  pivot_wider(names_from = symbol, values_from = adjusted)
```



```{r}
data_wide_mat = data_wide[,-1]
data_wide_mat = as.matrix(data_wide_mat)
cov_mat = cov(data_wide_mat, use = "pairwise.complete.obs")
```


Running `glasso`

```{r}
library(glasso)
glasso = glasso(cov_mat, rho = 3000000)
glasso_w = glasso$w
glasso_w_BTC = tickers[which(glasso_w != 0)]
glasso_w_BTC = glasso_w_BTC[!is.na(glasso_w_BTC)]
glasso_w_BTC
```
